A Pandemic-Scale Ancestral Recombination Graph for SARS-CoV-2

### 2025-10-08 latex/main.tex
Averted muts similar to ripples parsimony improvement
Closes #827

\subsection*{Quantification of evidence for recombination}
The LS HMM at the heart of sc2ts provides a simple and powerful means 
of quantifying the plausibility of a given recombination event.
To do this, we force the HMM to find a Viterbi solution for the 
putative recombinant sequence with no recombination by specifying
a large recombination penalty $k$ (Figure~\ref{fig:methodology}).
The difference between the numbers of mutations between
the two solutions---the number of mutations
``averted'' by recombination---is a measure of
the evidence in favor of recombination.
Table~\ref{tab:pango_x_lineages} shows the number of averted mutations
for each of the 21 recombination events associated with Pango X lineages,
with support ranging from very strong in the case of XC
(the most parsimonious non-recombinant placement would
require 37 additional mutations) to marginal for XG and XW,
which have the minimal possible number of averted mutations.

>>>>

\subsection*{Quantification of evidence for recombination}
The LS HMM at the heart of sc2ts provides a simple and powerful means 
of quantifying the plausibility of a given recombination event.
To do this, we force the HMM to find a Viterbi solution for the 
putative recombinant sequence with no recombination by specifying
a large recombination penalty $k$ (Figure~\ref{fig:methodology}).
The difference between the numbers of mutations between
the two solutions---the number of mutations
``averted'' by recombination---is a measure of
the evidence in favor of recombination  (similar to 
idea of ``parsimony improvement'' in RIPPLES.)
Table~\ref{tab:pango_x_lineages} shows the number of averted mutations
for each of the 21 recombination events associated with Pango X lineages,
with support ranging from very strong in the case of XC
(the most parsimonious non-recombinant placement would
require 37 additional mutations) to marginal for XG and XW,
which have the minimal possible number of averted mutations.
